Classification of samples by gene expression data
========================================================

Dr. Susanne Weller
29/04/2013

Import and transform data

```{r}
testpanel <- as.data.frame(read.table("HKPanel.csv", header=TRUE))
testpanel$Accession <- NULL
testpanel$Class_Name <- NULL
testpanel$Species_Name <- NULL
testpanel$Probe_Name <- NULL
testpanel$Avg_Count <- NULL
testpanel$Min_Count <- NULL
testpanel$Max_Count <- NULL
testpanel$X.CV <- NULL


# Transpose table YOU WANT
ttestpanel <- as.data.frame(t(testpanel[,2:ncol(testpanel)]))

# Set the column headings
colnames(ttestpanel) <- testpanel$Gene
ttestpanel$sample <- row.names(ttestpanel)
row.names(ttestpanel) <- NULL 
```

Unsupervised classification methods for gene expression data
------------------------------------------------------------

### Pricipal component analysis

```{r}
pca.out=prcomp((ttestpanel[1:32]),scale=TRUE)
pca.out
names(pca.out)
biplot(pca.out, scale=0)

```
### PCA Using FactoMinR

-This gives a much nicer plot
-dimdesc(pca) tells you the most important variables, which could be used to sort data to create a heat map that shows both clusters nicely.

```{r}
require(FactoMineR) 
# PCA with function PCA

pca <- PCA(ttestpanel[1:32], scale.unit=TRUE, ncp=5, graph=T)
#scale all the features,  ncp: number of dimensions kept in the results (by default 5)

gene_load <- dimdesc(pca)
gene_load
gene_importance <- as.data.frame(gene_load$Dim.1$quanti)
gene_importance$gene <-row.names(gene_importance)
row.names(gene_importance) <- NULL
#This line of code will sort the variables the most linked to each PC. It is very useful when you have many variables.
```

### k-means clustering
- this will always give a clustering an should only be used as a control!

```{r}

km.out=kmeans(ttestpanel[1:32],2,nstart=30)
km.out

```

Heatmap using the "heatmap"" function
-----------------------------------------
- This needs the genes to be the rownames!
```{r}
heattestpanel <- testpanel
rownames(heattestpanel) <- testpanel$Gene
heattestpanel$Gene <- NULL

heattestpanel = as.numeric(heattestpanel)
heatmap(as.matrix(heattestpanel,rownames.force=TRUE))
```

Heatmap using "gplot" package
```{r}
library("gplots")
heattestpanel = as.numeric(heattestpanel)
heatmap.2((as.matrix(heattestpanel, rownames.force=TRUE)), col=greenred(75), scale="row", key=TRUE, symkey=FALSE, density.info="none", trace="none", cexRow=0.5)